### 2023-07-31 Text/coverLetter.Rmd

---
output: pdf_document
#word_document

geometry: "left=3cm,right=3cm,top=1cm,bottom=1cm"

header-includes:
  - \usepackage{fontawesome5}
  - \pagenumbering{gobble} #for no page numbering
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=FALSE, warning = FALSE)
```

```{r}
journal <- "Peer Community In Ecology"
editor <- "Recommenders of PCI"
title <- "The role of seed-dispersing animals in shaping plant distributions"
listOfAuthors <- paste("Benjamin Robira", collapse = ", ")
```

\begin{flushleft}
\rule[-0.15in]{0.25\linewidth}{0.8ex}
\vspace{-0.8ex}
\hrule
\vspace{0.1in}
To the recommenders of \textit{`r journal`}
\end{flushleft}

\begin{flushright}
\faEnvelope \hspace{0.01in} \href{mailto:benjamin.robira@normalesup.org}{benjamin.robira@normalesup.org}\\
\faHome \hspace{0.01in} Université de Montpellier/CEFE/CNRS\\
\faMapPin \hspace{0.01in} 1919, route de Mende, 34293 Montpellier cedex 5\\

\hfill

On the `r Sys.Date()`,\\

\end{flushright}

\hfill

\begin{center}
Dear `r editor`,
\end{center}

\hfill

I hereby submit my manuscript entitled "`r title`" for consideration as a Research Article to *`r journal`*.

|   In the scenarios concerning the emergence and selection of spatiotemporal cognitive abilities in vagile plant-eating animals, there is always an implicit assumption: the distribution of plants does not change and ultimately shapes the cognitive abilities of the animals. Over the lifetime of an individual, the distribution of plants is likely to remain largely unchanged. However, over long periods of time, the resource landscape can change as a result of animal exploitation. In particular, animal movement can help shape the environment by dispersing seeds. Yet, animal movement is not unique and can be associated with a range of cognitive abilities, leading to a variety of movement patterns and hence possible resource patterns. Using a minimal set of assumptions in agent-based simulations, I show how animal foraging cognition can be both a cause and a consequence of resource heterogeneity. Specifically, I show how the emergence of routine behaviour is linked to the transformation of the resource landscape, with the materialisation of foraging routes through the deposition of seeds along them. In turn, this spatial engineering cascades into the selection gradient of cognitive abilities themselves, as the benefits of spatiotemporal memory are linked to the distribution of resources.

|   This study illustrates the possible existence of eco-evolutionary feedback loops between foragers' cognition and their environment. It therefore offers a new perspective on co-evolution between taxa, such as the angiosperm/primate dyad. I propose to contact for review Lisieux Fuzessy (CREAF, Universitat Autònoma de Barcelona, l.fuzessy@creaf.uab.cat), Chloe Bracis (Université Grenoble Alpes, Chloe.Bracis@univ-grenoble-alpes.fr), Merel Soons (Utrecht University, m.b.soons@uu.nl) and Joshua Plotkin (University of Pennsylvania, jplotkin@sas.upenn.edu) for their respective complementary expertise in zoochory and primate evolution, memory and animal movement modelling, and plant distribution (for the two latter). I have provided the links to the codes in the article to make them available for review. As this is only a code-based analysis, any error in the code could be fatal. A double check would therefore be greatly appreciated. I certify that this is not currently under submission elsewhere.

\hfill

\begin{flushright}
Sincerely yours,\\
Benjamin Robira.\\

\vspace{0.1in}
\includegraphics[width=0.55\textwidth]{signature.png}\\
\end{flushright}

<!-- Recommender: -->
<!-- Fabien Laroche -->
<!-- Erin Vogel -->
<!-- Aurelie Coulon -->
<!-- Luca Borger -->
<!-- Cedric Sueur -->